---
title: "FinalProjectJZeramby"
author: "Jessica Zeramby"
date: "6/30/2021"
output: rmarkdown::github_document
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

install.packages(c('tinytex', 'rmarkdown'))
tinytex::install_tinytex()

```{r, load data, include=FALSE}
library(tidyverse)
library(ggplot2)
library(tidyr)
library(plotly)
library(scales)
library(readr)
library(data.table)
library(lubridate)
library(flextable)
library(gridExtra)
library(maps)
library(mapdata)
library(viridis)
library(ggrepel)
library(dplyr)

#uploading raw data files as CSV files
X2014Records <- read_csv("2014Records.csv")
X2016Records <- read_csv("2016Records.csv")
X2018Records <- read_csv("2018Records.csv")
X2019Records <- read_csv("2019Records.csv")
Intakebytown_mappingdata <- read_csv("Top intake Towns WRARI csv.csv")
```

# BIO 539 Final Project: An Exploratory Data Analysis of Wildlife Rehabilitation Records in Southern Rhode Island

*Wildlife Rehabilitation Data: An Underutilized Resource for Characterizing Trends in Human-Wildlife Conflicts.*

## Introduction to Data
The Wildlife Rehabilitation Association of Rhode Island (WRARI) is the state's only fully operational, non-profit wildlife rehabilitation center and veterinary clinic. The WRARI staff granted permission for the following data to be used in this analysis. The WRARI staff willingly provided their annual wildlife intake records from 2014, 2016, 2018, and 2019 for the purposes of this analysis. These yearly intake records were provided in the form of spreadsheets containing data on every animal admitted to the facility in each year.

## Analysis of Wildlife Rehabilitation Records conducted to answer four key questions
1.	What are the most common causes of injury to wildlife in Rhode Island?
2.	Which species are most threatened or vulnerable in Rhode Island?
3.	How many animals are admitted to this rehabilitation facility on a regular basis, or what is the average number of intakes to this facility?
4.	Which Rhode Island towns serve as hotspots for injured or displaced wildlife?


*At the request of the data source, some sections of the following data files were omitted from this analysis. This type of data includes personal or private information as well as information that is not publicly accessible.*

## Start of Data Analysis
Starting with 2014 Intake Records

### 2014 
The first step in analyzing the 2014 intake records was determining the number of intakes in each month of the year.

```{r, echo = FALSE, error=TRUE}
#starting with 2014 data
#clean 2014 data
Records_14 <- rename(X2014Records, date = "Intake Date", species_class = "Species Class", species = "Species", specific_location = "Specific Location", city = "City Found", cause_for_admission = "Cause for Admission")
#extract important columns
Records_14 <- select(Records_14, c(1,2,3,7,8,9))

#change date format for 2014 
Records_14 <- Records_14 %>%
mutate(date = mdy(date))

#Add month column from date column for 2014
Records_14$month <- month(Records_14$date)

Records_14$month <- as.character(Records_14$month)

#remove NA for 2014 data
Records_14 <- Records_14 %>%
  filter(!is.na(date))

#plotting the intakes each month against months in the year 2014
updated_14_intakebymonth <- ggplot(Records_14) + aes(x = month(month, label = TRUE)) +
  geom_bar(fill = "#0099f9") +
  labs(x = "Month", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption =  element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Monthly Wildlife Admissions in 2014", caption = "Figure 1: The number of monthly admissions to the wildlife clinic in 2014.")
updated_14_intakebymonth 

#error occuring that did not previously occur prior to Rmarkdown process, "Error in month(month, label = TRUE) : unused argument (label = TRUE)" unsure why error is occuring now 
```
**Figure 1: Monthly admissions to the wildlife clinic in 2014.**

```{r, echo = FALSE}
#calculate total intakes per month in tibble for 2014 intakes
#monthly total caluclations for intakes admission 
Monthlystats_14 <- Records_14 %>%
  group_by(month) %>%
  tally()

#count monthly intakes in 2014 
Monthlystats_14 <- Records_14 %>%
  group_by(month) %>%
  summarise(count = n()) %>%
  na.omit("Monthlystats_14") %>%
  top_n(12)

#statistics of monthly 2014 admissions 
mean(Monthlystats_14$count)
#mean of monthly intakes = 177, meaning the average monthly intake in 2014 is 177 intakes per month

min(Monthlystats_14$count)
#minimum intake per month = 41, meaning the lowest number of monthly intakes in 2014 was 41 intakes in a month

max(Monthlystats_14$count)
#maximum intake per month = 458, meaning the maximum number of intakes in a month in 2014 was 458 intakes in a month

median(Monthlystats_14$count)
#median intake per month = 111, meaning of the range for monthly intakes in 2014 the middle number of intakes is 111 intakes in a month

#tibble of monthly intakes, convert numeric month format to abbreviated month 
Monthlystats_14$month <- (month.abb[Monthlystats_14$month])

#convert monthly stats 14 tibble into data frame
Monthlystats_14_table <- data.frame(Monthlystats_14)

#convert monthly stats 14 data frame into table
setDT(Monthlystats_14_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(Monthlystats_14_table)

#rename columns in data table
setnames(Monthlystats_14_table, c("month","count"), c("Month", "Admissions"))

#use flextable package
Monthlystats_14_tableflex <- flextable(Monthlystats_14_table)

#format table
Monthlystats_14_tableflex <- theme_vanilla(Monthlystats_14_tableflex)
Monthlystats_14_tableflex <- border_outer(Monthlystats_14_tableflex)
Monthlystats_14_tableflex <- border_inner_v(Monthlystats_14_tableflex)
Monthlystats_14_tableflex <- set_caption(Monthlystats_14_tableflex, "Table 1: The number of monthly admissions to the wildlife clinic in 2014.")

Monthlystats_14_tableflex
```
**Table 1: Monthly Admissions to the wildlife clinic in 2014**

According to these results, the average number of monthly intakes in 2014 was 177 total intakes per month. The month of June had the highest number of intakes in a single month, with 458 total intakes, followed by July, which had the second highest number of intakes per month for the year, with 374 total intakes. After peaking in the summer, the number of monthly intakes began to decline. Admissions fell steadily throughout October and November, reaching a winter low of 41 intakes in December, 56 in January, and 46 in February. According to these patterns, the number of wildlife admissions to this facility appears to follow seasonal trends.

The next step in analyzing the 2014 admissions data was to identify the top five reasons for admission to the clinic this year.
```{r, echo = FALSE}

#begin analysis for determining top reasons for admission

#check class for cause for admission column before replacing values
class(Records_14$cause_for_admission)

#replace values to fix spelling and grammar errors
#replace "orphaned" with "Orphaned" to match other variables of the same kind
Records_14$cause_for_admission <- gsub('orphaned', 'Orphaned', Records_14$cause_for_admission)

#replace "injured" with "Unkonwn trauma" to match other variables of the same kind
Records_14$cause_for_admission <- gsub('Injured', 'Unknown trauma', Records_14$cause_for_admission)

#top reasons for admission 2014 
admissions_2014 <- Records_14 %>%
  group_by(date, cause_for_admission) %>%
  tally()

#count frequency of reasons for admission in 2014
admissions_2014 <- na.omit("admissions_2014") 
admissions_2014_count <- Records_14 %>%
  group_by(cause_for_admission) %>%
  summarise(count = n())


#count top 5 common reasons for admission in 2014
admissions_2014 <- na.omit("admissions_2014") 
admissions_2014 <- Records_14 %>%
  group_by(cause_for_admission) %>%
  summarise(count = n()) %>%
  na.omit("admissions_2014") %>%
  top_n(5) 

#plot top 5 common reasons for admission in 2014
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels
top5_admissions14_plot <- ggplot(data = admissions_2014, aes(x = reorder(cause_for_admission, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#69b3a2") +
  labs(x = "Reason for Admission", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), 
        plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Five Reasons for Admission in 2014", caption = "Figure 2: The total number of admissions for each of the \n top five reasons for admission to the wildlife clinic in 2014.")
top5_admissions14_plot
```
**Figure 2: Total number of admissions for each of the top five reasons for admission to the wildlife clinic in 2014.** 

```{r, echo = FALSE}

#making the top5reasonsforadmission2014 in a table format
#make table with top 5 reasons, count, and percentage of total 

#arrange data frame into descending order
admissions_2014 <- admissions_2014 %>%
  arrange(desc(count))

#add percent to admissions14 table
admissions_2014 <- admissions_2014 %>%
  mutate(admissions_2014, Percent_total_intakes = count / sum (admissions_2014_count$count) * 100) 

#convert admissions 14 tibble into data frame
top5_rzn14_table <- data.frame(admissions_2014)

#convert admissions 14 data frame into table
setDT(top5_rzn14_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(top5_rzn14_table)

#rename columns in data table
setnames(top5_rzn14_table, c("cause_for_admission","count", "Percent_total_intakes"), 
         c("Reason Admitted", "Number of Admissions", "Percent of Total Admissions"))

#round percent column to whole numbers
top5_rzn14_table$`Percent of Total Admissions` <- round(top5_rzn14_table$`Percent of Total Admissions`)

#use flextable package
top5_rzn14_tableflex <- flextable(top5_rzn14_table)

#format table
#adjust dimensions to autofit
top5_rzn14_tableflex <- top5_rzn14_tableflex %>%
  autofit(add_w = 0, add_h = 0)

#format table
top5_rzn14_tableflex <- theme_vanilla(top5_rzn14_tableflex)
top5_rzn14_tableflex <- border_outer(top5_rzn14_tableflex)
top5_rzn14_tableflex <- set_caption(top5_rzn14_tableflex, "Table 2: The total number of admissions and percentage of total admissions for each of the top five reasons for admission in 2014.")
top5_rzn14_tableflex

#stats for top reasons for admission data (2014)
sum(top5_rzn14_table$`Number of Admissions`)
#top 5 reasons account for 1711 intakes of the 2133 total admissions for the year 2014
#top 5 reasons for intake account for 80% of all intakes that year
```
**Table 2: The total number of admissions and percentage of total admissions for each of the top five reasons for admission in 2014.**

In 2014, the top five reasons for admission to the wildlife clinic accounted for 1,711 of the 2,133 total intakes, or 80 percent of all intakes. Therefore, the top five reasons for admission accounted for more than three-quarters of all clinic admissions in 2014. The most common reason for admission in this year was orphaned wildlife (38%), followed by cases of unknown injuries or trauma (29%), attacks on wildlife by domestic cats (6%), incidents involving animal-vehicle collisions (4%), and attacks on wildlife by domestic dogs (2%). 

Following this segment of the 2014 data analysis, the species that were the most severely impacted and thus frequently admitted to the clinic in 2014 were evaluated.

```{r, echo = FALSE}
#begin analysis for top 5 most common species admitted in 2014 
#count top 5 most common species admitted in 2014

#top species admitted 2014 
topspecies_2014 <- Records_14 %>%
  group_by(species) %>%
  tally()

#count freq of all different species admitted in 2014
topspecies_2014 <- na.omit("topspecies_2014") 
topspecies_2014_count <- Records_14 %>%
  group_by(species) %>%
  summarise(count = n())

#count top 5 most common species admitted in 2014
topspecies_2014 <- na.omit("topspecies_2014") 
topspecies_2014 <- Records_14 %>%
  group_by(species) %>%
  summarise(count = n()) %>%
  na.omit("topspecies_2014") %>%
  top_n(5)
topspecies_2014

#plot top 5 species admitted in 2014
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels

top5_species14_plot <- ggplot(data = topspecies_2014, aes(x = reorder(species, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#CC3300") +
  labs(x = "Species", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5),  plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.8)) +
  theme(axis.text.x = element_text(size = 7, face = "bold", vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Five Most Frequently Admitted Species in 2014", caption = "Figure 3: The top five most frequently admitted species in 2014\n and the total number of clinic admissions for each.")
saveRDS(top5_species14_plot, file = "top5_species14_plot.RDS")
save(top5_species14_plot, file = "top5_species14_plot.RData")
top5_species14_plot
```
**Figure 3: The top five most frequently admitted species in 2014 and the total number of clinic admissions for each.**

```{r, echo=FALSE}
#making top 5 species table with top 5 species, number admissions (count), and percent total admissions

#arrange tibble into descending order
topspecies_2014 <- topspecies_2014 %>%
  arrange(desc(count))

#add percent to tibble
topspecies_2014 <- topspecies_2014 %>%
  mutate(topspecies_2014, 'Percent of Total Admissions' = count / sum (topspecies_2014_count$count) * 100) 
topspecies_2014

#convert top species tibble into data frame
top5_species14_table <- data.frame(topspecies_2014)

#convert admissions 14 data frame into table
setDT(top5_species14_table)


#check if this was sucessful in turning data frame into data table 
is.data.table(top5_species14_table)

#rename columns in data table
setnames(top5_species14_table, c("species","count", "Percent.of.Total.Admissions"), 
         c("Species", "Number of Admissions", "Percent of Total Admissions"))

#round percent column to whole numbers
top5_species14_table$`Percent of Total Admissions` <- round(top5_species14_table$`Percent of Total Admissions`)

#use flextable package
top5_species14_tableflex <- flextable(top5_species14_table)

#format table
#adjust dimensions to autofit
top5_species14_tableflex <- top5_species14_tableflex %>%
  autofit(add_w = 0, add_h = 0)

#format table
top5_species14_tableflex <- theme_vanilla(top5_species14_tableflex)
top5_species14_tableflex <- border_outer(top5_species14_tableflex)
top5_species14_tableflex <- set_caption(top5_species14_tableflex, "Table 3: The number of admissions and percent of total admissions for the top five most frequently admitted species in 2014.")
top5_species14_tableflex

#stats for top species (2014)

#sum of admissions from all top 5 common species combined
sum(top5_species14_table$'Number of Admissions')
#961 admissions
#961 of 2133 total admissions in 2014 are from top 5 species,
#top 5 species = 45% of all admissions in 2014

#compare admissions from top 5 species with total number of species admitted in 2014
Number_diffspecies_2014_forstats<- Records_14 %>%
  count(species)
#168
#168 different species admitted in 2014
#168 species total. Top 5 most commonly admitted species = almost 50% of all admissions
```
**Table 3: The number of admissions and percent of total admissions for the top five most frequently admitted species in 2014.**

According to these results, the top five most commonly admitted species accounted for 961 of the 2,133 total intakes in 2014, or 45 percent of all intakes in 2014. The Eastern Cotton Tail was the most frequently admitted species, accounting for 16% of all admissions, followed by the Eastern Grey Squirrel (10%), the House Sparrow (8%), the American Robin (8%), and the Rock Dove (3%). According to these findings, of the 168 different species admitted to the clinic in 2014, the top five most frequently admitted species accounted for nearly half of all admissions.

The portion of this analysis relating to the 2014 yearly intake records was completed after identifying the top five reasons for admission, top five species commonly admitted, and total monthly admissions for the year 2014. Following that, the 2016 intake records were evaluated in a similar manner.

### 2016 
To begin the 2016 intake record data analysis, the number of intakes in each month of the year was determined.

```{r, echo=FALSE, error=TRUE}
#Begin 2016 data analysis
#repeat steps above for 2016, clean, # of intakes, top 5 admission reasons, top 5 species admitted

#getinfoabout2016
summary(X2016Records)
X2016Records

#extract important columns only
Records_16<- select(X2016Records, c(1,2,3,8,9))

#clean data 2016
Records_16 <- rename(Records_16, date = "Intake Date", species_class = "Species Class", species = "Species", city = "City Found", cause_for_admission = "Cause for Admission")

#change date format for 2016 
Records_16 <- Records_16 %>%
mutate(date = mdy(date))

#Add month column from date column for 2016
Records_16$month <- month(Records_16$date)

#remove NA for 2016 data
Records_16 <- Records_16 %>%
  filter(!is.na(date))

#plotting the intakes each month against months in the year 2016
Intakebymonth_16 <- ggplot(Records_16) + aes(x = month(month, label = TRUE)) +
  geom_bar(fill = "#0099f9") +
  labs(x = "Month", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption =  element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Monthly Wildlife Admissions in 2016", caption = "Figure 4: The number of monthly admissions to the wildlife clinic in 2016.")
Intakebymonth_16

#unkown error occuring that was not occuring prior to Rmarkdown process, "Error in month(month, label = TRUE) : unused argument (label = TRUE)", unsure why error is occuring now 
```
**Figure 4: Monthly admissions to the wildlife clinic in 2016.**

```{r, echo=FALSE}
#2016 intake by month data into table 

#calculate total intakes per month in tibble for 2016 intakes
#monthly total calculations for intakes admission 
Monthlystats_16 <- Records_16 %>%
  group_by(month) %>%
  tally()

#count monthly intakes in 2016 
Monthlystats_16 <- Records_16 %>%
  group_by(month) %>%
  summarise(count = n()) %>%
  na.omit("Monthlystats_16") %>%
  top_n(12)

#statistics of monthly 2016 admissions 
mean(Monthlystats_16$count)
#mean of monthly intakes = 233, meaning the average monthly intake in 2016 is 233 intakes per month

min(Monthlystats_16$count)
#minimum intake per month = 38, meaning the lowest number of monthly intakes in 2016 was 38 intakes in a month

max(Monthlystats_16$count)
#maximum intake per month = 607, meaning the maxmimum number of intakes in a month in 2016 was 607 intakes in a month

median(Monthlystats_16$count)
#median intake per month = 185, meaning of the range for monthly intakes in 2016 the middle number of intakes is 185 intakes in a month


#tibble of monthly intakes, convert numeric month format to abbreviated month 
Monthlystats_16$month <- (month.abb[Monthlystats_16$month])

#convert monthly stats 16 tibble into data frame
Monthlystats_16_table <- data.frame(Monthlystats_16)

#convert monthly stats 16 data frame into table
setDT(Monthlystats_16_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(Monthlystats_16_table)

#rename columns in data table
setnames(Monthlystats_16_table, c("month","count"), c("Month", "Admissions"))

#use flextable package
Monthlystats_16_tableflex <- flextable(Monthlystats_16_table)

#format table
Monthlystats_16_tableflex <- theme_vanilla(Monthlystats_16_tableflex)
Monthlystats_16_tableflex <- border_outer(Monthlystats_16_tableflex)
Monthlystats_16_tableflex <- border_inner_v(Monthlystats_16_tableflex)
Monthlystats_16_tableflex <- set_caption(Monthlystats_16_tableflex, "Table 4: The number of monthly admissions to the wildlife clinic in 2016.")
Monthlystats_16_tableflex
```
**Table 4: Monthly Admissions to the wildlife clinic in 2016.**

In 2016, monthly admissions trends appeared to follow seasonal patterns similar to those observed in 2014. Again, clinic admissions peaked between May and August, with June and July continuing to be the busiest months for wildlife admissions, with 607 and 441 admissions per month, respectively. The highest number of admissions in a single month was 607, and the lowest number of monthly admissions was 38 total admissions. In a given month, the average number of total admissions was 233. From late summer to early fall, the number of admissions continued to fall until it reached its lowest point of the year in the winter months of December and January, with only 38 and 48 intakes, respectively.

In this section of the analysis, the next step was to determine the most common reasons for clinic admission in 2016.
```{r, echo=FALSE}
#begin analysis for determining top reasons for admission 2016

#check class for cause for admission column before replacing values
class(Records_16$cause_for_admission)

#replace values to fix spelling and grammar errors
#replace "Orphaned Juvenile" with "Orphaned" to match other variables of the same kind
Records_16$cause_for_admission <- gsub('Orphaned juvenile', 'Orphaned', Records_16$cause_for_admission)

#replace "injury" with "Unknown trauma" to match other variables of the same kind
Records_16$cause_for_admission <- gsub('Injury', 'Unknown trauma', Records_16$cause_for_admission)

#replace "HBC" with "Hit by car" to match other variables of the same kind
Records_16$cause_for_admission <- gsub('HBC', 'Hit by car', Records_16$cause_for_admission)

#replace "Cat caught" with "Cat attack" to match other variables of the same kind
Records_16$cause_for_admission <- gsub('Cat caught', 'Cat attack', Records_16$cause_for_admission)

#top reasons for admission 2016 
admissions_2016 <- Records_16 %>%
  group_by(date, cause_for_admission) %>%
  tally()

#count freq of reasons for admission in 2016
admissions_2016 <- na.omit("admissions_2016") 
admissions_2016_count <- Records_16 %>%
  group_by(cause_for_admission) %>%
  summarise(count = n())

#count top 5 common reasons for admission in 2016
admissions_2016 <- na.omit("admissions_2016") 
admissions_2016 <- Records_16 %>%
  group_by(cause_for_admission) %>%
  summarise(count = n()) %>%
  na.omit("admissions_2016") %>%
  top_n(5) 
admissions_2016

#plot top 5 common reasons for admission in 2016
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels
top5_admissions16_plot <- ggplot(data = admissions_2016, aes(x = reorder(cause_for_admission, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#69b3a2") +
  labs(x = "Reason for Admission", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Five Reasons for Admission in 2016", caption = "Figure 5: The total number of admissions for each of the \n top five reasons for admission to the wildlife clinic in 2016")
top5_admissions16_plot
```
**Figure 5: Total number of admissions for each of the top five reasons for admission to the wildlife clinic in 2016.**

```{r, echo=FALSE}
#making the top5reasonsforadmission2016 in a table format
#make table with top 5 reasons, count, and percentage of total 

#arrange data frame into descending order
admissions_2016 <- admissions_2016 %>%
  arrange(desc(count))
#add percent to admissions16 table
admissions_2016 <- admissions_2016 %>%
  mutate(admissions_2016, Percent_total_intakes = count / sum (admissions_2016_count$count) * 100) 
admissions_2016

#convert admissions 16 tibble into data frame
top5_rzn16_table <- data.frame(admissions_2016)

#convert admissions 16 data frame into table
setDT(top5_rzn16_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(top5_rzn16_table)

#rename columns in data table
setnames(top5_rzn16_table, c("cause_for_admission","count", "Percent_total_intakes"), 
         c("Reason Admitted", "Number of Admissions", "Percent of Total Admissions"))

#round percent column to whole numbers
top5_rzn16_table$`Percent of Total Admissions` <- round(top5_rzn16_table$`Percent of Total Admissions`)

#use flextable package
top5_rzn16_tableflex <- flextable(top5_rzn16_table)

#format table
#adjust dimensions to autofit
top5_rzn16_tableflex <- top5_rzn16_tableflex %>%
  autofit(add_w = 0, add_h = 0)

#format table
top5_rzn16_tableflex <- theme_vanilla(top5_rzn16_tableflex)
top5_rzn16_tableflex <- border_outer(top5_rzn16_tableflex)
top5_rzn16_tableflex<- set_caption(top5_rzn16_tableflex, "Table 5: The total number of admissions and percentage of total admissions for each of the top five reasons for admission in 2016.")

#stats for top reasons for admission data (2016)
sum(top5_rzn16_table$`Number of Admissions`)
#top 5 reasons account for 1064 intakes of the 2797 total admissions for the year 2016
#top 5 reasons for intake account for 38% of all intakes that year
top5_rzn16_tableflex
```
**Table 5: The total number of admissions and percentage of total admissions for each of the top five reasons for admission in 2016.**

In 2016, the top five primary reasons for admission to the wildlife clinic accounted for 1,064 of the 2,797 total intakes, or 38% of all intakes. According to these results, the most common reason for admission in 2016 was wildlife orphanage (27%), followed by cases of unknown injuries or trauma (5%), attacks on wildlife by domestic cats (3%), incidents involving animal-vehicle collisions (2%), and injuries to wildlife resulting in broken wings (1%).

The final step in the 2016 intake record data analysis was to determine the top five most common species admitted in 2016.

```{r, echo=FALSE}
#begin analysis for top 5 most common species admitted in 2016
#count top 5 most common species admitted in 2016

#top species admitted 2016 
topspecies_2016<- Records_16 %>%
  group_by(species) %>%
  tally()

#count freq of all different species admitted in 2016
topspecies_2016 <- na.omit("topspecies_2016") 
topspecies_2016_count <- Records_16 %>%
  group_by(species) %>%
  summarise(count = n())

#count top 5 most common species admitted in 2016
topspecies_2016 <- na.omit("topspecies_2016") 
topspecies_2016 <- Records_16 %>%
  group_by(species) %>%
  summarise(count = n()) %>%
  na.omit("topspecies_2016") %>%
  top_n(5)
topspecies_2016

#plot top 5 species admitted in 2016
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels

top5_species16_plot <- ggplot(data = topspecies_2016, aes(x = reorder(species, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#CC3300") +
  labs(x = "Species", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5),  plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.8)) +
  theme(axis.text.x = element_text(size = 7, face = "bold", vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Five Most Frequently Admitted Species in 2016", caption = "Figure 6: The top five most frequently admitted species in 2016\n and the total number of clinic admissions for each.")
saveRDS(top5_species16_plot, file = "top5_species16_plot.RDS")
save(top5_species16_plot, file = "top5_species16_plot.RData")
top5_species16_plot
```
**Figure 6: The top five most frequently admitted species in 2016 and the total number of clinic admissions for each.**

```{r, echo=FALSE}
#making top 5 species table with top 5 species, number admissions (count), and percent total admissions

#arrange tibble into descending order
topspecies_2016 <- topspecies_2016 %>%
  arrange(desc(count))

#add percent to tibble
topspecies_2016 <- topspecies_2016 %>%
  mutate(topspecies_2016, 'Percent of Total Admissions' = count / sum (topspecies_2016_count$count) * 100) 
topspecies_2016

#convert top species tibble into data frame
top5_species16_table <- data.frame(topspecies_2016)

#convert admissions 16 data frame into table
setDT(top5_species16_table)


#check if this was sucessful in turning data frame into data table 
is.data.table(top5_species16_table)

#rename columns in data table
setnames(top5_species16_table, c("species","count", "Percent.of.Total.Admissions"), 
         c("Species", "Number of Admissions", "Percent of Total Admissions"))

#round percent column to whole numbers
top5_species16_table$`Percent of Total Admissions` <- round(top5_species16_table$`Percent of Total Admissions`)

#use flextable package
top5_species16_tableflex <- flextable(top5_species16_table)

#format table
#adjust dimensions to autofit
top5_species16_tableflex <- top5_species16_tableflex %>%
  autofit(add_w = 0, add_h = 0)

#format table
top5_species16_tableflex <- theme_vanilla(top5_species16_tableflex)
top5_species16_tableflex <- border_outer(top5_species16_tableflex)
top5_species16_tableflex <- set_caption(top5_species16_tableflex, "Table 6: The number of admissions and percent of total admissions for the top five most frequently admitted species in 2016.")

#stats for top species (2016)

#sum of admissions from all top 5 common species combined
sum(top5_species16_table$'Number of Admissions')
#1069 admissions
#1069 of 2797 total admissions in 2016 are from top 5 species,
#top 5 species = 38% of all admissions in 2016

#compare admissions from top 5 species with total number of species admitted in 2016
Number_diffspecies_2016_forstats<- Records_16 %>%
  count(species)
Number_diffspecies_2016_forstats
#272
#272 different species admitted in 2016
#272 species total. Top 5 most commonly admitted species = almost 40% of all admissions

top5_species16_tableflex 
```
**Table 6: The number of admissions and percent of total admissions for the top five most frequently admitted species in 2016.**

According to these results, the top five most commonly admitted species accounted for 1,069 of the 2,797 total intakes, or almost 40 percent of all intakes in 2016. The Eastern Cotton Tail was once again the most frequently admitted species,  accounting for 13% of all admissions, followed by the Eastern Grey Squirrel (9%), the House Sparrow (7%), the American Robin (6%), and the Mourning Dove (3%). Out of the 272 different species admitted to the clinic in 2016, the top five most frequently admitted species accounted for nearly 40% of all admissions in this year.

Following the analysis of the 2016 admission records, the 2018 yearly intake records were next to be examined. By reviewing admission records from 2018, the number of monthly admissions to the clinic, the top five most common reasons for intake, and the most commonly admitted species in 2018 were all investigated again.

### 2018
The first step in analyzing the 2018 yearly intake record data was determining the number of monthly intakes in 2018.

```{r, echo=FALSE}
#Begin 2018 data analysis
#repeat steps above for 2018, clean, # of intakes, top 5 admission reasons, top 5 species admitted

#extract important columns only
Records_18<- select(X2018Records, c(1,2,3,8,9))

#clean data 2018
Records_18 <- rename(Records_18, date = "Intake Date", species_class = "Species Class", species = "Species", city = "City Found", cause_for_admission = "Cause for Admission")

#change date format for 2018
Records_18 <- Records_18 %>%
mutate(date = mdy(date))

#Add month column from date column for 2018
Records_18$month <- month(Records_18$date)

#remove NA for 2018 data
Records_18 <- Records_18 %>%
  filter(!is.na(date))

#plotting the intakes each month against months in the year 2018
Intakebymonth_18 <- ggplot(Records_18) + aes(x = month(month, label = TRUE)) +
  geom_bar(fill = "#0099f9") +
  labs(x = "Month", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption =  element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Monthly Wildlife Admissions in 2018", caption = "Figure 7: The number of monthly admissions to the wildlife clinic in 2018.")
Intakebymonth_18

#same error appearing for unknown reason, ran without error prior to Rmarkdown
#see previous monthly intake plots for further explanation
```
**Figure 7: Monthly admissions to the wildlife clinic in 2018.**

```{r, echo=FALSE}
#2018 intake by month data into table 

#calculate total intakes per month in tibble for 2018 intakes
#monthly total calculations for intakes admission 
Monthlystats_18 <- Records_18 %>%
  group_by(month) %>%
  tally()

#count monthly intakes in 2018 
Monthlystats_18 <- Records_18 %>%
  group_by(month) %>%
  summarise(count = n()) %>%
  na.omit("Monthlystats_18") %>%
  top_n(12)

#statistics of monthly 2018 admissions 

sum(Monthlystats_18$count)
#2626 total intakes in the year

mean(Monthlystats_18$count)
#mean of monthly intakes = 219, meaning the average monthly intake in 2018 is 219 intakes per month

min(Monthlystats_18$count)
#minimum intake per month = 48 , meaning the lowest number of monthly intakes in 2018 was 48  intakes in a month

max(Monthlystats_18$count)
#maximum intake per month = 507, meaning the maxmimum number of intakes in a month in 2018 was 507 intakes in a month

median(Monthlystats_18$count)
#median intake per month = 151, meaning of the range for monthly intakes in 2018 the middle number of intakes is 151 intakes in a month


#tibble of monthly intakes, convert numeric month format to abbreviated month 
Monthlystats_18$month <- (month.abb[Monthlystats_18$month])

#convert monthly stats 18 tibble into data frame
Monthlystats_18_table <- data.frame(Monthlystats_18)

#convert monthly stats 18 data frame into table
setDT(Monthlystats_18_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(Monthlystats_18_table)

#rename columns in data table
setnames(Monthlystats_18_table, c("month","count"), c("Month", "Admissions"))

#use flextable package
Monthlystats_18_tableflex <- flextable(Monthlystats_18_table)

#format table
Monthlystats_18_tableflex <- theme_vanilla(Monthlystats_18_tableflex)
Monthlystats_18_tableflex <- border_outer(Monthlystats_18_tableflex)
Monthlystats_18_tableflex <- border_inner_v(Monthlystats_18_tableflex)
Monthlystats_18_tableflex <- set_caption(Monthlystats_18_tableflex, "Table 7: The number of monthly admissions to the wildlife clinic in 2018.")
Monthlystats_18_tableflex
```
**Table 7: Monthly Admissions to the wildlife clinic in 2018.**

These findings indicate that monthly admissions trends in 2018 followed seasonal patterns as well. Admission numbers peaked again between May and August, with June being the busiest month for wildlife admissions, with 507 total admissions, and May and July coming in second with 414 and 412 intakes, respectively. June set the record for the most monthly admissions, while February had the fewest, with only 46 admissions in the entire month. Monthly admission numbers averaged 218 total admissions per month.

The next step in the 2018 data analysis was to evaluate the top reasons for clinic admission in 2018.

```{r, echo=FALSE}
#begin analysis for determining top reasons for admission 2018

#check class for cause for admission column before replacing values
class(Records_18$cause_for_admission)

#replace values to fix spelling and grammar errors
#replace "Orphaned Juvenile" with "Orphaned" to match other variables of the same kind
Records_18$cause_for_admission <- gsub('Orphaned juvenile', 'Orphaned', Records_18$cause_for_admission)

#replace "injury" with "Unknown trauma" to match other variables of the same kind
Records_18$cause_for_admission <- gsub('Injury', 'Unknown trauma', Records_18$cause_for_admission)

#replace "HBC" with "Hit by car" to match other variables of the same kind
Records_18$cause_for_admission <- gsub('HBC', 'Hit by car', Records_18$cause_for_admission)

#replace "Cat caught" with "Cat attack" to match other variables of the same kind
Records_18$cause_for_admission <- gsub('Cat caught', 'Cat attack', Records_18$cause_for_admission)

#top reasons for admission 2018
admissions_2018 <- Records_18 %>%
  group_by(date, cause_for_admission) %>%
  tally()

#count freq of reasons for admission in 2018
admissions_2018 <- na.omit("admissions_2018") 
admissions_2018_count <- Records_18 %>%
  group_by(cause_for_admission) %>%
  summarise(count = n())

#count top 5 common reasons for admission in 2018
admissions_2018 <- na.omit("admissions_2018") 
admissions_2018 <- Records_18 %>%
  group_by(cause_for_admission) %>%
  summarise(count = n()) %>%
  na.omit("admissions_2018") %>%
  top_n(5) 
admissions_2018

#plot top 5 common reasons for admission in 2018
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels
top5_admissions18_plot <- ggplot(data = admissions_2018, aes(x = reorder(cause_for_admission, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#69b3a2") +
  labs(x = "Reason for Admission", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Five Reasons for Admission in 2018", caption = "Figure 8: The total number of admissions for each of the \n top five reasons for admission to the wildlife clinic in 2018")
top5_admissions18_plot
```
**Figure 8: Total number of admissions for each of the top five reasons for admission to the wildlife clinic in 2018.** 

```{r, echo=FALSE}
#making the top5reasonsforadmission2018 in a table format
#make table with top 5 reasons, count, and percentage of total 

#arrange data frame into descending order
admissions_2018 <- admissions_2018 %>%
  arrange(desc(count))
#add percent to admissions18 table
admissions_2018 <- admissions_2018 %>%
  mutate(admissions_2018, Percent_total_intakes = count / sum (admissions_2018_count$count) * 100) 
admissions_2018

#convert admissions 18 tibble into data frame
top5_rzn18_table <- data.frame(admissions_2018)

#convert admissions 18 data frame into table
setDT(top5_rzn18_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(top5_rzn18_table)

#rename columns in data table
setnames(top5_rzn18_table, c("cause_for_admission","count", "Percent_total_intakes"), 
         c("Reason Admitted", "Number of Admissions", "Percent of Total Admissions"))

#round percent column to whole numbers
top5_rzn18_table$`Percent of Total Admissions` <- round(top5_rzn18_table$`Percent of Total Admissions`)

#use flextable package
top5_rzn18_tableflex <- flextable(top5_rzn18_table)

#format table
#adjust dimensions to autofit
top5_rzn18_tableflex <- top5_rzn18_tableflex %>%
  autofit(add_w = 0, add_h = 0)

#format table
top5_rzn18_tableflex <- theme_vanilla(top5_rzn18_tableflex)
top5_rzn18_tableflex <- border_outer(top5_rzn18_tableflex)
top5_rzn18_tableflex<- set_caption(top5_rzn18_tableflex, "Table 8: The total number of admissions and percentage of total admissions for each of the top five reasons for admission in 2018.")

#stats for top reasons for admission data (2018)
sum(top5_rzn18_table$`Number of Admissions`)
#784
#top 5 reasons account for 784  intakes of the 2626 total admissions for the year 2018
#top 5 reasons for intake account for 29% of all intakes in 2018

top5_rzn18_tableflex
```
**Table 8: The total number of admissions and percentage of total admissions for each of the top five reasons for admission in 2018.**

According to these findings, the top five primary reasons for admission to the wildlife clinic accounted for 784 of the 2,626 total intakes in 2018, or 29 percent of all intakes. The most common reason for admission in 2018 was wildlife orphanage (20%), followed by cases of unknown injuries or trauma (5%), attacks on wildlife by domestic cats (2%), injuries to wildlife resulting in broken wings (1%), and incidents involving animal-vehicle collisions (1%). 

Following these findings, the next step in the 2018 analysis was to determine which species were the most severely impacted and thus frequently admitted to the clinic this year.

```{r, echo=FALSE}
#begin analysis for top 5 most common species admitted in 2018
#count top 5 most common species admitted in 2018

#top species admitted 2018
topspecies_2018<- Records_18 %>%
  group_by(species) %>%
  tally()

#count freq of all different species admitted in 2018
topspecies_2018 <- na.omit("topspecies_2018") 
topspecies_2018_count <- Records_18 %>%
  group_by(species) %>%
  summarise(count = n())

#count top 5 most common species admitted in 2018
topspecies_2018 <- na.omit("topspecies_2018") 
topspecies_2018 <- Records_18 %>%
  group_by(species) %>%
  summarise(count = n()) %>%
  na.omit("topspecies_2018") %>%
  top_n(5)
topspecies_2018

#plot top 5 species admitted in 2018
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels

top5_species18_plot <- ggplot(data = topspecies_2018, aes(x = reorder(species, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#CC3300") +
  labs(x = "Species", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5),  plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.8)) +
  theme(axis.text.x = element_text(size = 7, face = "bold", vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Five Most Frequently Admitted Species in 2018", caption = "Figure 9: The top five most frequently admitted species in 2018\n and the total number of clinic admissions for each.")
top5_species18_plot
```
**Figure 9: The top five most frequently admitted species in 2018 and the total number of clinic admissions for each.**

```{r, echo=FALSE}
#making top 5 species table with top 5 species, number admissions (count), and percent total admissions

#arrange tibble into descending order
topspecies_2018 <- topspecies_2018 %>%
  arrange(desc(count))

#add percent to tibble
topspecies_2018 <- topspecies_2018 %>%
  mutate(topspecies_2018, 'Percent of Total Admissions' = count / sum (topspecies_2018_count$count) * 100) 
topspecies_2018

#convert top species tibble into data frame
top5_species18_table <- data.frame(topspecies_2018)

#convert admissions 18 data frame into table
setDT(top5_species18_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(top5_species18_table)

#rename columns in data table
setnames(top5_species18_table, c("species","count", "Percent.of.Total.Admissions"), 
         c("Species", "Number of Admissions", "Percent of Total Admissions"))

#round percent column to whole numbers
top5_species18_table$`Percent of Total Admissions` <- round(top5_species18_table$`Percent of Total Admissions`)

#use flextable package
top5_species18_tableflex <- flextable(top5_species18_table)

#format table
#adjust dimensions to autofit
top5_species18_tableflex <- top5_species18_tableflex %>%
  autofit(add_w = 0, add_h = 0)

#format table
top5_species18_tableflex <- theme_vanilla(top5_species18_tableflex)
top5_species18_tableflex <- border_outer(top5_species18_tableflex)
top5_species18_tableflex <- set_caption(top5_species18_tableflex, "Table 9: The number of admissions and percent of total admissions for the top five most frequently admitted species in 2018.")
top5_species18_tableflex

#stats for top species (2018)

#sum of admissions from all top 5 common species combined
sum(top5_species18_table$'Number of Admissions')
#1186 admissions
#1186 of 2626 total admissions in 2018 are from top 5 species,
#top 5 species = 45% of all admissions in 2018

#compare admissions from top 5 species with total number of species admitted in 2018
Number_diffspecies_2018_forstats<- Records_18 %>%
  count(species)
Number_diffspecies_2018_forstats
#199
#199 different species admitted in 2018
#199 species total. Top 5 most commonly admitted species = almost 45% of all admissions

top5_species18_tableflex
```
**Table 9: The number of admissions and percent of total admissions for the top five most frequently admitted species in 2018.**

According to these findings, the top five most frequently admitted species accounted for 1,186 of the total 2,626 admissions in 2018, or 45 percent of all admissions. This year, the Eastern Cotton Tail was the most frequently admitted species, accounting for 17% of all admissions, followed by the Eastern Grey Squirrel (11%), the House Sparrow (7%), the American Robin (6%), and the Herring Gull (4%). As a result, of the 199 species admitted to the clinic in 2018, admission from the top five most frequently admitted species accounted for 45 percent of all admissions.

The intake records from 2019 were the next and final set to be examined in this analysis. As previously stated, these records were analyzed to determine the total number of monthly admissions to the clinic, the top five most common reasons for intake, and the most commonly admitted species.

### 2019

Finally, the intake records for 2019 were the final set of yearly intake records to be examined in this analysis. The first step in analyzing the 2019 intake data was determining the number of monthly admissions in each month of the year.

```{r, echo=FALSE}
#Begin 2019 data analysis
#repeat steps above for 2019, clean, # of intakes, top 5 admission reasons, top 5 species admitted

#extract important columns only
Records_19<- select(X2019Records, c(1,2,3,8,9))

#clean data 2019
Records_19 <- rename(Records_19, date = "Intake Date", species_class = "Species Class", species = "Species", city = "City Found", cause_for_admission = "Cause for Admission")

#change date format for 2019
Records_19 <- Records_19 %>%
mutate(date = mdy(date))

#Add month column from date column for 2019
Records_19$month <- month(Records_19$date)

#remove NA for 2019 data
Records_19 <- Records_19 %>%
  filter(!is.na(date))

#plotting the intakes each month against months in the year 2019
Intakebymonth_19 <- ggplot(Records_19) + aes(x = month(month, label = TRUE)) +
  geom_bar(fill = "#0099f9") +
  labs(x = "Month", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption =  element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Monthly Wildlife Admissions in 2019", caption = "Figure 10: The number of monthly admissions to the wildlife clinic in 2019.")
Intakebymonth_19
```
**Figure 10: Monthly admissions to the wildlife clinic in 2019.**

```{r, echo=FALSE}
#2019 intake by month data into table 

#calculate total intakes per month in tibble for 2019 intakes
#monthly total calculations for intakes admission 
Monthlystats_19 <- Records_19 %>%
  group_by(month) %>%
  tally()

#count monthly intakes in 2019
Monthlystats_19 <- Records_19 %>%
  group_by(month) %>%
  summarise(count = n()) %>%
  na.omit("Monthlystats_19") %>%
  top_n(12)

#statistics of monthly 2019 admissions 

sum(Monthlystats_19$count)
#3014 total intakes in the year

mean(Monthlystats_19$count)
#mean of monthly intakes = 251, meaning the average monthly intake in 2019 is 251 intakes per month

min(Monthlystats_19$count)
#minimum intake per month = 62 , meaning the lowest number of monthly intakes in 2019 was 62  intakes in a month

max(Monthlystats_19$count)
#maximum intake per month = 537, meaning the maximum number of intakes in a month in 2019 was 537 intakes in a month

median(Monthlystats_19$count)
#median intake per month =184, meaning of the range for monthly intakes in 2019 the middle number of intakes is 184 intakes in a month

#tibble of monthly intakes, convert numeric month format to abbreviated month 
Monthlystats_19$month <- (month.abb[Monthlystats_19$month])

#convert monthly stats 19 tibble into data frame
Monthlystats_19_table <- data.frame(Monthlystats_19)

#convert monthly stats 19 data frame into table
setDT(Monthlystats_19_table)

#check if this was successful in turning data frame into data table 
is.data.table(Monthlystats_19_table)

#rename columns in data table
setnames(Monthlystats_19_table, c("month","count"), c("Month", "Admissions"))

#use flextable package
Monthlystats_19_tableflex <- flextable(Monthlystats_19_table)

#format table
Monthlystats_19_tableflex <- theme_vanilla(Monthlystats_19_tableflex)
Monthlystats_19_tableflex <- border_outer(Monthlystats_19_tableflex)
Monthlystats_19_tableflex <- border_inner_v(Monthlystats_19_tableflex)
Monthlystats_19_tableflex <- set_caption(Monthlystats_19_tableflex, "Table 10: The number of monthly admissions to the wildlife clinic in 2019.")
Monthlystats_19_tableflex
```
**Table 10: Monthly Admissions to the wildlife clinic in 2019.**

According to these results, monthly admissions in 2019 followed similar seasonal patterns as in previous years. Admission numbers peaked between May and August once again, with June being the busiest month for wildlife admissions, followed by July, which had the second highest admission count for the year. The highest monthly total number of admissions was 537 in June, and the lowest monthly total number of admissions was 62 in February. The average monthly admissions totaled 251 total admissions per month. The number of intakes remained close to its peak throughout the summer, with intakes falling just below 500 in August.

Next, the top five reasons for clinic admission in 2019 were evaluated.

```{r, echo=FALSE}
#check class for cause for admission column before replacing values
class(Records_19$cause_for_admission)

#replace values to fix spelling and grammar errors
#replace "Orphaned Juvenile" with "Orphaned" to match other variables of the same kind
Records_19$cause_for_admission <- gsub('Orphaned juvenile', 'Orphaned', Records_19$cause_for_admission)

#replace "injury" with "Unknown trauma" to match other variables of the same kind
Records_19$cause_for_admission <- gsub('Injury', 'Unknown trauma', Records_19$cause_for_admission)

#replace "HBC" with "Hit by car" to match other variables of the same kind
Records_19$cause_for_admission <- gsub('HBC', 'Hit by car', Records_19$cause_for_admission)

#replace "Cat caught" with "Cat attack" to match other variables of the same kind
Records_19$cause_for_admission <- gsub('Cat caught', 'Cat attack', Records_19$cause_for_admission)

Records_19$cause_for_admission <- gsub('Possible HBC', 'Hit by car', Records_19$cause_for_admission)
Records_19$cause_for_admission <- gsub('Unknown', 'Unknown trauma', Records_19$cause_for_admission)
Records_19$cause_for_admission <- gsub('FX wing', 'Broken wing', Records_19$cause_for_admission)
Records_19$cause_for_admission <- gsub('Back injury', 'Unknown trauma', Records_19$cause_for_admission)

Records_19$cause_for_admission [grepl("Cat", Records_19$cause_for_admission)] <- "Cat attack"

#top reasons for admission 2019
admissions_2019 <- Records_19 %>%
  group_by(date, cause_for_admission) %>%
  tally()


#count freq of reasons for admission in 2019
admissions_2019 <- na.omit("admissions_2019") 
admissions_2019_count <- Records_19 %>%
  group_by(cause_for_admission) %>%
  summarise(count = n())

#count top 5 common reasons for admission in 2019
admissions_2019 <- na.omit("admissions_2019") 
admissions_2019 <- Records_18 %>%
  group_by(cause_for_admission) %>%
  summarise(count = n()) %>%
  na.omit("admissions_2019") %>%
  top_n(5) 

#plot top 5 common reasons for admission in 2019
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels
top5_admissions19_plot <- ggplot(data = admissions_2019, aes(x = reorder(cause_for_admission, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#69b3a2") +
  labs(x = "Reason for Admission", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Five Reasons for Admission in 2019", caption = "Figure 11: The total number of admissions for each of the \n top five reasons for admission to the wildlife clinic in 2019.")
top5_admissions19_plot
```
**Figure 11: Total number of admissions for each of the top five reasons for admission to the wildlife clinic in 2019.**

```{r, echo=FALSE}
#making the top5reasonsforadmission2019 in a table format
#make table with top 5 reasons, count, and percentage of total 

#arrange data frame into descending order
admissions_2019 <- admissions_2019 %>%
  arrange(desc(count))

#add percent to admissions19 table
admissions_2019 <- admissions_2019 %>%
  mutate(admissions_2019, Percent_total_intakes = count / sum (admissions_2019_count$count) * 100) 
admissions_2019

#convert admissions 19 tibble into data frame
top5_rzn19_table <- data.frame(admissions_2019)

#convert admissions 19 data frame into table
setDT(top5_rzn19_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(top5_rzn19_table)

#rename columns in data table
setnames(top5_rzn19_table, c("cause_for_admission","count", "Percent_total_intakes"), 
         c("Reason Admitted", "Number of Admissions", "Percent of Total Admissions"))

#round percent column to whole numbers
top5_rzn19_table$`Percent of Total Admissions` <- round(top5_rzn19_table$`Percent of Total Admissions`)

#use flextable package
top5_rzn19_tableflex <- flextable(top5_rzn19_table)

#format table
#adjust dimensions to autofit
top5_rzn19_tableflex <- top5_rzn19_tableflex %>%
  autofit(add_w = 0, add_h = 0)

#format table
top5_rzn19_tableflex <- theme_vanilla(top5_rzn19_tableflex)
top5_rzn19_tableflex <- border_outer(top5_rzn19_tableflex)
top5_rzn19_tableflex <- set_caption(top5_rzn19_tableflex, "Table 11: The total number of admissions and percentage of total admissions for each of the top five reasons for admission in 2019.")
top5_rzn19_tableflex

#stats for top reasons for admission data (2019)
sum(top5_rzn19_table$`Number of Admissions`)
#784
#top 5 reasons account for 784 intakes of the 3014 total admissions for the year 2019
#top 5 reasons for intake account for 26% of all intakes in 2019
top5_rzn19_tableflex
```
**Table 11: The total number of admissions and percentage of total admissions for each of the top five reasons for admission in 2019.**

According to these findings, the top five primary reasons for admission to the wildlife clinic accounted for 784 of the 3014 total intakes in 2019, or 26 percent of all intakes. The most common reason for admission in 2019 was wildlife orphanage (17%), followed by cases of unknown injuries or trauma (5%), domestic cat attacks on wildlife (2%), wildlife injuries resulting in broken wings (1%), and incidents involving animal-vehicle collisions (1%)

Following this section of the 2019 data analysis, the species that were the most severely impacted and thus frequently admitted to the clinic in 2019 were assessed.
```{r, echo=FALSE}
#begin analysis for top 5 most common species admitted in 2019
#count top 5 most common species admitted in 2019

#top species admitted 2019
topspecies_2019<- Records_19 %>%
  group_by(species) %>%
  tally()

#count freq of all different species admitted in 2019
topspecies_2019 <- na.omit("topspecies_2019") 
topspecies_2019_count <- Records_19 %>%
  group_by(species) %>%
  summarise(count = n())

#count top 5 most common species admitted in 2019
topspecies_2019 <- na.omit("topspecies_2019") 
topspecies_2019 <- Records_19 %>%
  group_by(species) %>%
  summarise(count = n()) %>%
  na.omit("topspecies_2019") %>%
  top_n(5)
topspecies_2019

#plot top 5 species admitted in 2019
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels

top5_species19_plot <- ggplot(data = topspecies_2019, aes(x = reorder(species, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#CC3300") +
  labs(x = "Species", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5),  plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.8)) +
  theme(axis.text.x = element_text(size = 7, face = "bold", vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Five Most Frequently Admitted Species in 2019", caption = "Figure 12: The top five most frequently admitted species in 2019\n and the total number of clinic admissions for each.")
top5_species19_plot
```
**Figure 12: The top five most frequently admitted species in 2019 and the total number of clinic admissions for each.**

```{r, echo=FALSE}
#making top 5 species table with top 5 species, number admissions (count), and percent total admissions

#arrange tibble into descending order
topspecies_2019 <- topspecies_2019 %>%
  arrange(desc(count))

#add percent to tibble
topspecies_2019 <- topspecies_2019 %>%
  mutate(topspecies_2019, 'Percent of Total Admissions' = count / sum (topspecies_2019_count$count) * 100) 
topspecies_2019

#convert top species tibble into data frame
top5_species19_table <- data.frame(topspecies_2019)

#convert admissions 19 data frame into table
setDT(top5_species19_table)


#check if this was successful in turning data frame into data table 
is.data.table(top5_species19_table)

#rename columns in data table
setnames(top5_species19_table, c("species","count", "Percent.of.Total.Admissions"), 
         c("Species", "Number of Admissions", "Percent of Total Admissions"))

#round percent column to whole numbers
top5_species19_table$`Percent of Total Admissions` <- round(top5_species19_table$`Percent of Total Admissions`)

#use flextable package
top5_species19_tableflex <- flextable(top5_species19_table)

#format table
#adjust dimensions to autofit
top5_species19_tableflex <- top5_species19_tableflex %>%
  autofit(add_w = 0, add_h = 0)

#format table
top5_species19_tableflex <- theme_vanilla(top5_species19_tableflex)
top5_species19_tableflex <- border_outer(top5_species19_tableflex)
top5_species19_tableflex <- set_caption(top5_species19_tableflex, "Table 12: The number of admissions and percent of total admissions for the top five most frequently admitted species in 2019.")

#stats for top species (2019)

#sum of admissions from all top 5 common species combined
sum(top5_species19_table$'Number of Admissions')
#1400 admissions
#1400 of 3014 total admissions in 2019 are from top 5 species,
#top 5 species = 46% of all admissions in 2019

#compare admissions from top 5 species with total number of species admitted in 2019
Number_diffspecies_2019_forstats<- Records_19 %>%
  count(species)
Number_diffspecies_2019_forstats
#169
#169 different species admitted in 2019
#169 species total. Top 5 most commonly admitted species = almost 50% of all admissions
top5_species19_tableflex
```
**Table 12: The number of admissions and percent of total admissions for the top five most frequently admitted species in 2019.**

In 2019, the top five most commonly admitted species accounted for 1,400 of the total 3,014 admissions, or 46 percent of all admissions. The Eastern Cotton Tail was the most frequently admitted species this year, accounting for 18% of all admissions, followed by the Eastern Grey Squirrel (12%), American Robin (6%), House Sparrow (5%), and Herring Gull (5%). According to these results, of the 169 different species admitted to the clinic in 2019, the top five most frequently admitted species accounted for 46 percent of all admissions overall.

Following the completion of the preceding analyses, the portion of this analysis pertaining to individual yearly intake record data from 2014, 2016, 2018, and 2019 was completed. Following that, the results of each of the yearly intake record data analyses were compared side by side to visualize changes in the data sets from one year to the next, with the goal of examining these intake records for larger picture trends spanning over longer time periods.

### Side-by-Side Analysis of Yearly Intake Record Data

To begin, the plots generated from each of the yearly intake record analyses for the number of monthly intakes, top reasons for admission, and top species admitted in each year were placed side by side to gain perspective on how these trends change over time.

```{r echo=FALSE, error=TRUE}
#start side by side comparison for all four years
#compare intakes monthly, top species, and top reasons for admission between all 4 years

#monthly intakes all years
Monthlyintakes_allyears <- grid.arrange(updated_14_intakebymonth, Intakebymonth_16, Intakebymonth_18, Intakebymonth_19, nrow = 2, top = "Monthly Admissions by Year")
Monthlyintakes_allyears
#same unknown error- see intake plots above for further details
```
**Figure 13: Monthly Admissions by Year, 2014-2018.**

According to these results, June had the highest number of wildlife admissions per month in each of the years for which data was available during this study period. Across all years of data collected during this study period, one recurring theme was a peak in clinic admissions from May to August each year, followed by a decline from early fall to winter, and a consistent increase in admission numbers from year to year when compared to previous years. While the continuous increase in admission numbers may be attributed to the wildlife clinic receiving more publicity and recognition from the local community, it is also possible that this is an indication of the declining health and well-being of the state's wildlife populations and overall ecosystem status.

```{r, echo=FALSE}
#top rzns admission all years
Top5rzns_allyears <- grid.arrange(top5_admissions14_plot, top5_admissions16_plot, top5_admissions18_plot, top5_admissions19_plot, nrow = 2, top = "Top Five Reasons for Admission by Year")
Top5rzns_allyears
```
**Figure 14: Top Five Reasons for Admission by Year, 2014-2018.**

According to these findings, the top reasons for admission to the clinic across all years of data included animals being orphaned as the leading cause and unknown trauma as the second leading cause. Cat attacks, broken wings, and car-animal collisions were the next most common reasons for admission. However, the number of admissions due to cat attacks, broken wings, and vehicle-animal collisions varied slightly from year to year in terms of which was more common than the others.

```{r, echo=FALSE}
#top species all years
Top5species_allyears <- grid.arrange(top5_species14_plot, top5_species16_plot, top5_species18_plot, top5_species19_plot, nrow = 2, top = "Top Five Common Species by year")
Top5species_allyears
```
**Figure 15: Top Five Common Species by Year, 2014-2018.**

In each of the years studied, the Eastern Cottontail, Eastern Grey Squirrel, House Sparrow, and American Robin were consistently among the top four of the top five most common species admitted to the clinic. The Eastern cottontail was the most common species admitted to the clinic in all four years, followed by the Eastern grey squirrel.

### Merged Data Analysis
Finally, in order to analyze the intake records from all of the years included in this analysis, all of the intake record files from each individual year were merged into a single dataset.

#stopped here Thurs 7.21


First, the merged yearly intake records were assessed for the number of monthly admissions over the entire four year study period.

```{r, echo=FALSE, error=TRUE}
#merging data
merged_intake <- merge(Records_14, Records_16, by=c ("month"))
merged_intake2 <- merge(Records_18, Records_19, by=c ("month"))
merge <- left_join(merged_intake, merged_intake2)

#plotting the intakes each year
Intakebymonth_all <- ggplot(merge) + aes(x = month(month, label = TRUE)) +
  geom_bar(fill = "#0099f9") +
  labs(x = "Month", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption =  element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  scale_y_continuous(label = label_number_si()) +
  labs(title = "Combined Monthly Wildlife Admissions, 2014-2019", caption = "Figure 16: The total number of admissions to the wildlife clinic by month, 2014-2018.")
Intakebymonth_all
```
**Figure 16: Total Wildlife Clinic Admissions by Month, 2014-2018.**

```{r, echo=FALSE, error=TRUE}

#2014-2018 intake by month data 

#calculate total intakes per month in tibble for all years intakes
#monthly total calculations for intakes admission 
Monthlystats_all <- merge %>%
  group_by(month) %>%
  tally()

#count monthly intakes all years
Monthlystats_all <- merge %>%
  group_by(month) %>%
  summarise(count = n()) %>%
  na.omit("Monthlystats_all") %>%
  top_n(12)

#tibble of monthly intakes, convert numeric month format to abbreviated month 
Monthlystats_all$month <- (month.abb[Monthlystats_all$month])

#convert monthly stats tibble into data frame
Monthlystats_all_table <- data.frame(Monthlystats_all)

#convert monthly stats data frame into table
setDT(Monthlystats_all_table)

#check if this was sucessful in turning data frame into data table 
is.data.table(Monthlystats_all_table)

#rename columns in data table
setnames(Monthlystats_all_table, c("month","count"), c("Month", "Admissions"))

#use flextable package
Monthlystats_all_tableflex <- flextable(Monthlystats_all_table)

#format table
Monthlystats_all_tableflex <- theme_vanilla(Monthlystats_all_tableflex)
Monthlystats_all_tableflex <- border_outer(Monthlystats_all_tableflex)
Monthlystats_all_tableflex <- border_inner_v(Monthlystats_all_tableflex)
Monthlystats_all_tableflex <- set_caption(Monthlystats_all_tableflex, "Table 13: The number of monthly admissions to the wildlife clinic by date, 2014-2018.")

#statistics of monthly admissions across all years
sum(Monthlystats_all$count)
#804332 total intakes across the years 2014, 2016, 2018, 2019

mean(Monthlystats_all$count)
#mean of monthly intakes = 67027, meaning across these four years of data the wildlife rehab averaged 67027 intakes

min(Monthlystats_all$count)
#minimum intake per month = 1558 , meaning the month with the least intakes per month across all data was December with a combined 1558 intakes from all four years

max(Monthlystats_all$count)
#maximum intake per month = 278006, meaning the month with the most intakes per month across all data was June, with 278006 total intakes per month combined across all four years

median(Monthlystats_all$count)
#median intake per month = 21875, meaning of the range for monthly intakes across all years of data or the middle number of intakes was 21875 intakes in a month
#statistics of monthly admissions across all years

sum(Monthlystats_all$count)
#804332 total intakes across the years 2014, 2016, 2018, 2019

Monthlystats_all_tableflex
```
**Table 13: Monthly Wildlife Clinic Admissions by Date, 2014-2018.**

The combined number of intakes per month remained highest across all four years of data in June, with a total of 278,006 total intakes recorded across all four years of data. From May to August, the combined admissions remained above 100,000, but dropped to the lowest combined number of intakes in December, accounting for only 1,558 total combined admissions across all four years of admission records.

The merged intake record files were then analyzed to determine the number of intakes year after year.

```{r, echo=FALSE}

#rbind all records data to make next plots
#remove extra column in Rec14
Records_14 <- select(Records_14, c(1,2,3,5,6,7))

#rbind data
Allyearsdata <- rbind(Records_14, Records_16, Records_18, Records_19)
#Add Year column to all data combined 
Allyearsdata$Year <- as.character(year(Allyearsdata$date))
Allyearsdata <- select(Allyearsdata, c(7))
#count observations in each year
Allyearsdata <- Allyearsdata %>%
  filter(Year != 2000) %>%
  filter(Year != 2001) %>%
  filter(Year != 2013) %>%
  filter(Year != 2015) %>%
  filter(Year != 2017)
Allyearsdata %>%
  count(Year)

Intakebyyear <- ggplot(Allyearsdata) + aes(x = Year) +
  geom_bar(fill = "#0099f9") +
  labs(x = "Year", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption =  element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Yearly Wildlife Admissions, 2014-2019", caption = "Figure 17: The number of yearly admissions 
       to the wildlife clinic from 2014 to 2019.")
Intakebyyear
```
**Figure 17: Yearly Admissions to the Clinic, 2014-2019.**

Between 2014, when there were a total of 2,133 admissions, and 2019, when there were a total of 3,014 admissions, the number of admissions to the wildlife clinic increased by nearly a thousand intakes. Although admission numbers fell slightly in 2018, from 2,797 in the previous year on record, 2016, to 2,626 in 2018, they continued to rise in 2019 and exceeded the number of intakes recorded in all previous years. These trends indicate that the number of wildlife admissions, and thus the number of injured and displaced wild animals, is steadily increasing year after year.

The top reasons for admissions across all years of data were the next to be evaluated from the merged data set.

```{r, echo=FALSE}
#all years top reasons 
#rbind data
Allyearsdata1 <- rbind(Records_14, Records_16, Records_18, Records_19)

#Add Year column to all data combined 
Allyearsdata1$Year <- as.character(year(Allyearsdata1$date))

#count observations in each year
Allyearsdata1 <- Allyearsdata1 %>%
  filter(Year != 2000) %>%
  filter(Year != 2001) %>%
  filter(Year != 2013) %>%
  filter(Year != 2015) %>%
  filter(Year != 2017)

#top reasons for admission all years
merge_admissions <- Allyearsdata1 %>%
  group_by(date, cause_for_admission) %>%
  tally()

#count top 5 common reasons for admission in all years
merge_admissions <- merge_admissions %>%
  group_by(cause_for_admission) %>%
  summarise(count = n()) %>%
  na.omit("merge_admissions") %>%
  top_n(5)

#plot top 5 common reasons for admission in 2019
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels
Merge_admissions_plot <- ggplot(data = merge_admissions, aes(x = reorder(cause_for_admission, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#69b3a2") +
  labs(x = "Reason for Admission", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
labs(title = "Top Five Reasons for Admission, by Date 2014-2019", caption = "Figure 18: The total number of admissions for each of the \n top five reasons for admission to the wildlife clinic, by date from 2014- 2019.")
Merge_admissions_plot
```
**Figure 18: The total number of admissions and percentage of total admissions for each of the top five reasons for admission, by date from 2014-2019.**

Following this, the top species admitted across all years of data in the study were examined.

```{r, echo=FALSE}
#all years top species
#rbind data
Allyearsdata2 <- rbind(Records_14, Records_16, Records_18, Records_19)
#replace values to fix spelling and grammar errors
#replace "Eastern Grey Squirrel" with "Eastern Gray Squirrel" to match other variables of the same kind
Allyearsdata2$cause_for_admission <- gsub('Eastern Grey Squirrel', 'Eastern Gray Squirrel', Allyearsdata2$cause_for_admission)
Allyearsdata2$cause_for_admission [grepl("Eastern Grey Squirrel", Allyearsdata2$cause_for_admission)] <- "Eastern Gray Squirrel"
#Add Year column to all data combined 
Allyearsdata2$Year <- as.character(year(Allyearsdata2$date))

#count observations in each year
Allyearsdata2 <- Allyearsdata2 %>%
  filter(Year != 2000) %>%
  filter(Year != 2001) %>%
  filter(Year != 2013) %>%
  filter(Year != 2015) %>%
  filter(Year != 2017)

Allyearsdata2$cause_for_admission [grepl("Eastern Grey Squirrel", Allyearsdata2$cause_for_admission)] <- "Eastern Gray Squirrel"

#top species all years
merge_species <- Allyearsdata2 %>%
  group_by(species) %>%
  tally()

#count top 5 common species in all years
merge_species <- Allyearsdata2 %>%
  group_by(species) %>%
  summarise(count = n()) %>%
  na.omit("merge_species") %>%
  top_n(3)

#plot top 5 common reasons for admission all years
#plot in descending order on x axis
#\n seperates label or title into next line
#added formatting code to axis, title, caption and labels
merge_species_plot <- ggplot(data = merge_species, aes(x = reorder(species, -count), y = count)) + 
  geom_bar(stat = "identity", fill = "#CC3300") +
  labs(x = "Species", y = "Number of Admissions") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold", margin = margin(6, 6, 6, 6), hjust = 0.5), plot.caption = element_text(face = "bold", hjust = 0.5, vjust = -0.6)) +
  theme(axis.text.x = element_text(size = 9, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.title.x = element_text(vjust =-0.55)) +
  theme(axis.title.y = element_text(vjust =0.55)) +
  labs(title = "Top Three Most Frequently Admitted Species, 2014 - 2019", caption = "Figure 19: The top three most frequently admitted species and the total \n number of clinic admissions for each, 2014- 2019.")
merge_species_plot
```

#stoppped here, add MAP, ADD RESULTS, update after merged and then done!
